\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-20 Documentation/TearingModeTheory.tex
Started Section 8

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\
Institute for Fusion St udies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\date{}
\maketitle

@@

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. (By definition, $a\ll R_0$ in a large aspect-ratio tokamak.( It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In most situations of interest, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

\section{Reduced Drift-MHD Resonant  Response Model}
\subsection{Introduction}
Equations~(\ref{e202})--(\ref{e206}) contain two quantities---namely, the reconnection time, $\tau_s$, and the unperturbed
natural frequency, $\omega_{s\,0}$---that can only be determined by solving for the plasma response in the radially-thin resonant layer surrounding the resonant surface. Unfortunately, determining the resonant plasma response is not a straightforward task. The
problem is that the small characteristic lengthscale in the resonant region (i.e., the resonant layer width) causes the fundamental ordering parameters
$\delta_e$ and $\delta_i$ (see Section~\ref{s3.9}) to take values that are of order unity. Consequently, most of the simplifications described in
Section~\ref{lowest} are not applicable in the resonant layer. In particular, many of the terms in the Braginskii equations (see Section~\ref{sbrag}) that are negligible in the outer region (e.g., plasma resistivity) have to be taken into account in the resonant layer.

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of a compressible-Alfv\'{e}n wave from them altogether, in the process converting a drift-magnetohydrodynamical (MHD) response model
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, et al.\ 1985; Furuya, et al.\ 2003). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf b}\right] = \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel$ and $\eta_\perp=1/\sigma_\perp$ are the parallel and perpendicular plasma
resistivities,  respectively. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$. Incidentally, if we assume that all
flow velocities are similar in magnitude to the diamagnetic velocity then the neglect of
electron inertia and electron viscosity in the previous equation is justified because the corresponding terms are factors
$(\rho_e/r_s)^2$ and $({\mit\Omega}_e\,\tau_e)\,(\rho_e/r_s)^2$, respectively, smaller than the leading order terms
in the equation.

@@

\subsection{Reduction Process} 
All variables in the resonant layer are assumed to be functions of 
$\hat{x}=(r-r_s)/l$,
 and 
 \begin{equation}
 \zeta=m\,\theta-n\,\varphi
 \end{equation}
 only. 
Let ${\bf n} = (0,\,\epsilon/q_s,\,1)$, where $\epsilon=r/R_0$ and 
\begin{equation}
q_s=q(r_s)= \frac{m}{n}.
\end{equation}
 It follows that ${\bf n}\cdot\nabla A=0$ for any $A(\hat{x},\zeta)$. 
Let us write (Fitzpatrick \& Waelbroeck 2009)
\begin{equation}\label{e238}
\hat{\bf B} \simeq (1+ \delta b)\,{\bf n} + \hat{\nabla}\hat{\psi}\times {\bf n},
\end{equation}
which automatically satisfies Equation~(\ref{e230}). (Recall that we are assuming that $|r-r_s|\ll r_s$ inside the resonant layer.) Note that $\hat{\psi}=\psi/(r_s\,B_z)$, where $\psi$ is defined in Equation~(\ref{e93}).
 Here, $\delta b(\hat{x},\zeta)$ and $\hat{\psi}(\hat{x},\zeta)$ are assumed to be small
compared to unity, which merely indicates that the perturbed magnetic field due to the tearing mode is small compared to the
equilibrium magnetic field.  We can write
\begin{equation}
\hat{\bf E} \simeq \left(\hat{E}_\parallel- \frac{\partial \hat{\psi}}{\partial \hat{t}}\right)\!{\bf n} + \hat{\nabla}\!\left(\frac{\partial\chi}{\partial \hat{t}}\right)\times {\bf n} + \hat{\nabla}\phi,
\end{equation}
which automatically satisfies Equation~(\ref{e231}) provided that
\begin{equation}
\hat{\nabla}^2\chi =\delta b.
\end{equation}
Here, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, where $E_\parallel$ represents the constant inductive electric field that drives the equilibrium plasma current in the vicinity of the resonant surface. Moreover, $\partial/\partial \hat{t}$ and $\phi(\hat{x},\zeta)$
are 
assumed to be first order in small quantities, whereas $\hat{E}_\parallel$ is assumed to be second order. 
It follows that
\begin{align}
\hat{\bf V}_E &\simeq \hat{\nabla}\phi\times {\bf n},\\[0.5ex]
\hat{\bf V} &\simeq \hat{V}_{\parallel\,i}\,{\bf n}+  \hat{\nabla}\phi\times {\bf n}+\hat{\nabla}{\mit\Xi},
\end{align}
where $\hat{V}_{\parallel\,i}(\hat{x},\zeta)$ is assumed to be first order in small quantities, whereas ${\mit\Xi}(\hat{x},\zeta)$ is assumed to be
second order. 
We can write
\begin{equation}
\hat{p}= \frac{\beta}{{\mit\Gamma}} + \delta p,
\end{equation}
where 
\begin{equation}\label{e244}
\beta= {\mit\Gamma}\,\hat{p}_0= \frac{{\mit\Gamma}\,p_0\,\mu_0}{B_z^{\,2}}.
\end{equation}
 Here, $\delta p(\hat{x},\zeta)$ is assumed to be first order in small
quantities. 
It follows that
\begin{align}
\hat{\bf V}_\ast &\simeq \hat{d}_i\,{\bf n}\times \hat{\nabla} \delta p,\\[0.5ex]
\hat{\bf V}_i&\simeq  \hat{V}_{\parallel\,i} \,{\bf n}+ \hat{\nabla}\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right)\times {\bf n}.
\end{align}

@@

We can write
\begin{align}
\hat{\nabla}\cdot \hat{\bf q} &=- \left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right)[[\delta p,\hat{\psi}],\hat{\psi}]\nonumber\\[0.5ex]
&\phantom{=} - \left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right)\nabla_\perp^{\,2} \delta p,\label{e260}
\end{align}
where $\hat{\nabla}_\perp^{\,2}\equiv \hat{\nabla}\cdot\hat{\nabla}_\perp$, $\chi_{\parallel\,e,i} = \kappa_{\parallel\,e,i}/n_0$, $\chi_{\perp\,e,i} = \kappa_{\perp\,e,i}/n_0$, $\hat{\chi}_{\parallel\,e,i}=
\chi_{\parallel\,e,i}/(l\,V_A)$,  $\hat{\chi}_{\perp\,e,i}=
\chi_{\perp\,e,i}/(l\,V_A)$, and use has been made of Equations~(\ref{e3.94a}),  (\ref{e3.94b}), (\ref{e211})--(\ref{e213}), and (\ref{e238}). 
We can also
write
\begin{equation}
\hat{\nabla}\cdot\hat{\bpi}_i =- \hat{\chi}_\varphi\,\hat{\nabla}^2\hat{\bf V}_i,
\end{equation}
where $\chi_\varphi = \mu_{\perp}/(n_0\,m_i)$, $\hat{\chi}_\varphi = \chi_\varphi/(l\,V_A)$, and $\mu_\perp$ is the anomalous
ion perpendicular viscosity. Note that the previous equation is entirely phenomenological in nature.  It is easily seen that
\begin{align}\label{e262}
{\bf n}\cdot\hat{\nabla}\cdot\hat{\bpi}_i&= -\hat{\chi}_\varphi\,\hat{\nabla}^{2}\hat{V}_{\parallel\,i},\\[0.5ex]
{\bf n}\cdot\hat{\nabla}\times \hat{\nabla}\cdot\hat{\bpi}_i&= \hat{\chi}_\varphi \,\hat{\nabla}^4\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right).\label{e263}
\end{align}
(Here, we are neglecting any $\hat{x}$ dependence of the various $\chi$ parameters for ease of notation.)

Let us define $V=\hat{V}_{\parallel\,i}$ and $N=\delta p/c_\beta$. Equations~(\ref{eJ}), (\ref{e252})--(\ref{e254}), (\ref{e256}), (\ref{e258})--(\ref{e260}), (\ref{e262}), and (\ref{e263}) yield the following closed set of equations (Fitzpatrick \& Waelbroeck 2005):
\begin{align}\label{e264}
\frac{\partial\hat{\psi}}{\partial\hat{t}}&= [\phi,\hat{\psi}] + \hat{d}_\beta\!\left(\frac{\tau}{1+\tau}\right)\left(1+\frac{0.71\,\eta_e}{1+\eta_e}\right)[N,\hat{\psi}]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial U}{\partial \hat{t}}&= [\phi,U] + \frac{\hat{d}_\beta}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\hat{\psi}] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi - \frac{\hat{d}_\beta\,N}{1+\tau}\right),\label{e265} \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] -c_\beta\,[V,\hat{\psi}] -\hat{d}_\beta\,[J,\hat{\psi}] + \hat{D}_\parallel\,[[N,\hat{\psi}],\hat{\psi}]
+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\label{e266}\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] - c_\beta\,[N,\hat{\psi}] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V,\label{e267}\\[0.5ex]
J &=-\frac{2}{q_s\,\hat{R}_0}+\hat{\nabla}^2\hat{\psi},\\[0.5ex]
U &=\hat{\nabla}^2 \phi.\label{e269}
\end{align}
Here, 
\begin{align}\label{e270}
d_\beta &= c_\beta\,d_i = \left(\frac{{\mit\Gamma}}{1+\beta}\right)^{1/2}\rho_s,\\[0.5ex]
 \rho_s &= \frac{v_t}{{\mit\Omega}_i},\\[0.5ex]
v_t &= \left(\frac{T_e+T_i}{m_i}\right)^{1/2},
\end{align}
and $\hat{d}_\beta=d_\beta/l$. The quantity $\rho_s$ is usually referred to as the {\em ion sound radius}. 
Moreover,
\begin{align}
\hat{D}_\parallel&= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp &= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right).
\end{align}

@@

Strictly speaking, the four-field fluid model 
presented in this section is only valid when the ion gyroradius is smaller than the typical variation lengthscale in the resonant layer. 
If this is not the case then it is necessary to adopt a so-called {\em gyrofluid}\/ approach in which the ion response is
averaged over the ion gyro-orbits (Cowley, Kulsrud \& Hahm 1986; Hammett, Dorland \& Perkins 1992; Waelbroeck, Hazeltine \& Morrison 2009). Of course, if the ions are cold (i.e., $T_i\ll T_e$) then the ion gyroradius tends to zero,  which means that the
analysis is valid. 
Note, however, that the ion sound radius is finite even when the ions are cold. In the following, we shall assume that our model remains valid even when the ions are hot.

>>>>

\begin{document}
\title{\bf Tearing Mode Dynamics in Tokamak Plasmas}
\author{R.~Fitzpatrick\,\footnote{rfitzp@utexas.edu}\\
Institute for Fusion Studies,\\  Department of Physics,\\  University of Texas at Austin,\\  Austin TX 78712, USA}
\date{}
\maketitle

@@

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. (By definition, $a\ll R_0$ in a large aspect-ratio tokamak.) It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In most situations of interest, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

\section{Reduced Drift-MHD Resonant  Response Model}\label{ff}
\subsection{Introduction}
Equations~(\ref{e202})--(\ref{e206}) contain two quantities---namely, the reconnection time, $\tau_s$, and the unperturbed
natural frequency, $\omega_{s\,0}$---that can only be determined by solving for the plasma response in the radially-thin resonant layer surrounding the resonant surface. Unfortunately, determining the resonant plasma response is not a straightforward task. The
problem is that the small characteristic lengthscale in the resonant region (i.e., the resonant layer width) causes the fundamental ordering parameters
$\delta_e$ and $\delta_i$ (see Section~\ref{s3.9}) to take values that are of order unity. Consequently, most of the simplifications described in
Section~\ref{lowest} are not applicable in the resonant layer. In particular, many of the terms in the Braginskii equations (see Section~\ref{sbrag}) that are negligible in the outer region (e.g., plasma resistivity) have to be taken into account in the resonant layer.

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of  compressible-Alfv\'{e}n waves from them altogether, in the process converting a drift-magnetohydrodynamical (MHD) response model
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, Kotschenreuther \& Morrison 1985; 
Furuya, Yagi \& Itoh 2003). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf b}\right] = \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel(r_s)$ and $\eta_\perp=1/\sigma_\perp(r_s)$ are the parallel and perpendicular plasma
resistivities,  respectively, at the resonant surface. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$. Incidentally, if we assume that all
flow velocities are similar in magnitude to the diamagnetic velocity then the neglect of
electron inertia and electron viscosity in the previous equation is justified because the corresponding terms are factors
$(\rho_e/r_s)^2$ and $({\mit\Omega}_e\,\tau_e)\,(\rho_e/r_s)^2$, respectively, smaller than the leading order terms
in the equation.

@@

\subsection{Reduction Process} 
All variables in the resonant layer are assumed to be functions of 
$\hat{x}=(r-r_s)/l$,
 and 
 \begin{equation}
 \zeta=m\,\theta-n\,\varphi
 \end{equation}
 only. 
Let ${\bf n} = (0,\,\epsilon/q_s,\,1)$, where $\epsilon=r/R_0$ and 
\begin{equation}
q_s=q(r_s)= \frac{m}{n}.
\end{equation}
 It follows that ${\bf n}\cdot\nabla A=0$ for any $A(\hat{x},\zeta)$. 
Let us write (Fitzpatrick \& Waelbroeck 2009)
\begin{equation}\label{e238}
\hat{\bf B} \simeq (1+ \delta b)\,{\bf n} + \hat{\nabla}\hat{\psi}\times {\bf n},
\end{equation}
which automatically satisfies Equation~(\ref{e230}). Note that $\hat{\psi}=\psi/(r_s\,B_z)$, where $\psi$ is defined in Equation~(\ref{e93}).
 Here, $\delta b(\hat{x},\zeta)$ and $\hat{\psi}(\hat{x},\zeta)$ are assumed to be small
compared to unity, which merely indicates that the perturbed magnetic field due to the tearing mode is small compared to the
equilibrium magnetic field.  We can write
\begin{equation}
\hat{\bf E} \simeq \left(\hat{E}_\parallel- \frac{\partial \hat{\psi}}{\partial \hat{t}}\right)\!{\bf n} + \hat{\nabla}\!\left(\frac{\partial\chi}{\partial \hat{t}}\right)\times {\bf n} + \hat{\nabla}\phi,
\end{equation}
which automatically satisfies Equation~(\ref{e231}) provided that
\begin{equation}
\hat{\nabla}^2\chi =\delta b.
\end{equation}
Here, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, where $E_\parallel$ represents the constant inductive electric field that drives the equilibrium plasma current in the vicinity of the resonant surface. Moreover, $\partial/\partial \hat{t}$ and $\phi(\hat{x},\zeta)$
are 
assumed to be first order in small quantities, whereas $\hat{E}_\parallel$ is assumed to be second order. 
It follows that
\begin{align}
\hat{\bf V}_E &\simeq \hat{\nabla}\phi\times {\bf n},\\[0.5ex]
\hat{\bf V} &\simeq \hat{V}_{\parallel\,i}\,{\bf n}+  \hat{\nabla}\phi\times {\bf n}+\hat{\nabla}{\mit\Xi},
\end{align}
where $\hat{V}_{\parallel\,i}(\hat{x},\zeta)$ is assumed to be first order in small quantities, whereas ${\mit\Xi}(\hat{x},\zeta)$ is assumed to be
second order. 
We can write
\begin{equation}
\hat{p}= \frac{\beta}{{\mit\Gamma}} + \delta p,
\end{equation}
where 
\begin{equation}\label{e244}
\beta= {\mit\Gamma}\,\hat{p}_0= \frac{{\mit\Gamma}\,p_0\,\mu_0}{B_z^{\,2}}.
\end{equation}
 Here, $\delta p(\hat{x},\zeta)$ is assumed to be first order in small
quantities. 
It follows that
\begin{align}
\hat{\bf V}_\ast &\simeq \hat{d}_i\,{\bf n}\times \hat{\nabla} \delta p,\\[0.5ex]
\hat{\bf V}_i&\simeq  \hat{V}_{\parallel\,i} \,{\bf n}+ \hat{\nabla}\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right)\times {\bf n}.
\end{align}

@@

We can write
\begin{align}
\hat{\nabla}\cdot \hat{\bf q} &=- \left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right)[[\delta p,\hat{\psi}],\hat{\psi}]\nonumber\\[0.5ex]
&\phantom{=} - \left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right)\nabla_\perp^{\,2} \delta p,\label{e260}
\end{align}
where $\hat{\nabla}_\perp^{\,2}\equiv \hat{\nabla}\cdot\hat{\nabla}_\perp$, $\chi_{\parallel\,e,i} = \kappa_{\parallel\,e,i}(r_s)/n_0$, $\chi_{\perp\,e,i} = \kappa_{\perp\,e,i}(r_s)/n_0$, $\hat{\chi}_{\parallel\,e,i}=
\chi_{\parallel\,e,i}/(l\,V_A)$,  $\hat{\chi}_{\perp\,e,i}=
\chi_{\perp\,e,i}/(l\,V_A)$, and use has been made of Equations~(\ref{e3.94a}),  (\ref{e3.94b}), (\ref{e211})--(\ref{e213}), and (\ref{e238}). 
We can also
write
\begin{equation}
\hat{\nabla}\cdot\hat{\bpi}_i =- \hat{\chi}_\varphi\,\hat{\nabla}^2\hat{\bf V}_i,
\end{equation}
where $\chi_\varphi = \mu_{\perp}(r_s)/(n_0\,m_i)$, $\hat{\chi}_\varphi = \chi_\varphi/(l\,V_A)$, and $\mu_\perp(r)$ is the anomalous
ion perpendicular viscosity. Note that the previous equation is entirely phenomenological in nature.  It is easily seen that
\begin{align}\label{e262}
{\bf n}\cdot\hat{\nabla}\cdot\hat{\bpi}_i&= -\hat{\chi}_\varphi\,\hat{\nabla}^{2}\hat{V}_{\parallel\,i},\\[0.5ex]
{\bf n}\cdot\hat{\nabla}\times \hat{\nabla}\cdot\hat{\bpi}_i&= \hat{\chi}_\varphi \,\hat{\nabla}^4\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right).\label{e263}
\end{align}

\subsection{Four-Field Model}\label{fourf}
Let us define $V=\hat{d}_i\,\hat{V}_{\parallel\,i}$ and $N=-\hat{d}_i\,\delta p$. Equations~(\ref{eJ}), (\ref{e252})--(\ref{e254}), (\ref{e256}), (\ref{e258})--(\ref{e260}), (\ref{e262}), and (\ref{e263}) yield the following closed set of equations (Fitzpatrick \& Waelbroeck 2005):
\begin{align}\label{e264}
\frac{\partial\hat{\psi}}{\partial\hat{t}}&= [\phi,\hat{\psi}] -\left(\frac{\tau}{1+\tau}\right)(1+\lambda_e)\,[N,\hat{\psi}]
+\hat{\eta}_\parallel\,J + \hat{E}_\parallel,\\[0.5ex]
\frac{\partial U}{\partial \hat{t}}&= [\phi,U] - \frac{1}{2\,(1+\tau)}\left(\hat{\nabla}^2[\phi,N] + [\hat{\nabla}^2\phi,N] + [\hat{\nabla}^2 N,\phi]\right) + [J,\hat{\psi}] \nonumber\\[0.5ex]&\phantom{=}+\hat{\chi}_\varphi  \,\hat{\nabla}^4\!\left(\phi + \frac{N}{1+\tau}\right),\label{e265} \\[0.5ex]
\frac{\partial N}{\partial \hat{t}}&= [\phi,N] +c_\beta^{\,2}\,[V,\hat{\psi}] +\hat{d}_\beta^{\,2}\,[J,\hat{\psi}] + \hat{D}_\parallel\,[[N,\hat{\psi}],\hat{\psi}]
+ \hat{D}_\perp\,\hat{\nabla}_\perp^{\,2}N,\label{e266}\\[0.5ex]
\frac{\partial V}{\partial\hat{t}}&= [\phi,V] +[N,\hat{\psi}] + \hat{\chi}_\varphi\,\hat{\nabla}^2 V,\label{e267}\\[0.5ex]
J &=-\frac{2}{q_s\,\hat{R}_0}+\hat{\nabla}^2\hat{\psi},\\[0.5ex]
U &=\hat{\nabla}^2 \phi.\label{e269}
\end{align}
Here, the quantity 
\begin{equation}\label{e270}
d_\beta = c_\beta\,d_i 
\end{equation}
 is usually referred to as the {\em ion sound radius}, and $\hat{d}_\beta =d_\beta/l$. 
Moreover,
\begin{equation}
 \lambda_e = 0.71\left(\frac{\eta_e}{1+\eta_e}\right),
\end{equation}
and
\begin{align}
\hat{D}_\parallel&=\frac{D_{\parallel}}{l\,V_A}= \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\parallel\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\parallel\,i}\right),\\[0.5ex]
\hat{D}_\perp &=\frac{D_{\perp}}{l\,V_A}= c_\beta^{\,2}\,\hat{\eta}_\perp +  \frac{2}{3}\left(1-c_\beta^{\,2}\right)\!\left(\frac{\eta_e}{1+\eta_e}\,\frac{\tau}{1+\tau}\,\hat{\chi}_{\perp\,e} + \frac{\eta_i}{1+\eta_i}\,\frac{1}{1+\tau}\,\hat{\chi}_{\perp\,i}\right).
\end{align}

@@

Strictly speaking, the four-field fluid model 
presented in this section is only valid when the ion gyroradius is smaller than the typical variation lengthscale in the resonant layer. 
If this is not the case then it is necessary to adopt a so-called {\em gyrofluid}\/ approach in which the ion response is
averaged over the ion gyro-orbits (Cowley, Kulsrud \& Hahm 1986; Hammett, Dorland \& Perkins 1992; Waelbroeck, Hazeltine \& Morrison 2009).
